<h5>Problem 2</h5>
(a)
```{r load_packages, include=FALSE, echo=FALSE}
library(compiler)
library(ggplot2)
```
```{r}
#step = number of iterations
sampler <- function(step) {
  #initialize intial graph
  M <- matrix(0,100,100)
  sums <- 0
  for(i in 1:step) {
    x <- trunc(runif(1,1,101)) #choose an x value at random
    y <- trunc(runif(1,1,101)) #choose a y value at random
    change <- trunc(runif(1,0,2)) #determine 1 try to flip to 1, or 0, flip to 0
    if(change) {
      #Now we will check to see if we can flip our new point
      flip <- TRUE
      if(x+1 <= 100) {
        if(M[x+1,y]==1) {flip <- FALSE} #Check right
      }
      if (x >=2 & flip) {
        if(M[x-1,y]==1) {flip <- FALSE} #Check left
      }
      if(y+1<=100 & flip) {
        if(M[x,y+1]==1) {flip<-FALSE} #Check above
      }
      if(y>=2 & flip) {
        if(M[x,y-1] == 1) {flip<-FALSE} #Check below
      }
      if(flip) {M[x,y]<-1}
    } else { M[x,y]<-0 }
  sums <- rbind(sums,sum(M)/10000)
  }
  return(sums)
}
```

<p>Now run the sampler 50,000 times and graph to find the steady state.</p>

```{r echo = FALSE}
M2 <- sampler(50000)
qplot(y = as.numeric(M2)) + xlab("Iteration") + ylab("Percent Occupied")
```

<p>It appears to reach a steady state around 20,000 - 30,000 iterations at roughly 20 percent occupied, so I will run the sampler 1000 times for 25,000 steps per chain and generate a histogram.</p>

```{r echo=FALSE}
M3 <- sapply(1:100, function(x) sampler(25000)) 
qplot(as.numeric(M3), geom="histogram", binwidth=.005) + xlab("f(x)/10000")
```

<p>(b) We can see from the previous graphs that the chain reaches a steady state around .23, so the probability that it will reach .4 is near zero.</p>

```{r}
montecarlo <- function(p, matrix) {
  M <- matrix #using the matrix we generated for the histogram for part a
  results <- 0
  for (i in 1:100) {
    p <- ifelse(M[i] > .4, 1, 0)
    results <- c(results, p)
  }
  return(sum(results)/100)
}
```

```{r echo=FALSE}
p <- montecarlo(.4, M3)
ci.right <- p + 2 * sqrt((p*(1-p)/100))
ci.left <- p - 2 * sqrt((p*(1-p)/100))
```

<p>Doing Monte Carlo integration yields a p-value of `r p` for the null hypothesis that e was drawn uniformly from all possible configurations.</p>

(c) 
```{r}
sampler2 <- function(step) {
  #initialize intial graph
  M <- matrix(0,100,100)
  M[50,50] <- 1 #because we can't divide by 0
  sums <- 0 #place to store sums of each step
  #now we need to come up with a valid proposal and the test whether we accept or reject based on whether a uniform rv < nu(j)/nu(i)
  for(i in 1:step) {
    base <- M
    #step 1: find a valid proposal
    x <- trunc(runif(1,1,101)) #choose an x coordinate
    y <- trunc(runif(1,1,101)) #choose a y value at random
    change <- trunc(runif(1,0,2)) #determine 1 try to flip to 1, or 0, flip to 0
    if(change) {
      flip <- TRUE #Now we will check to see if we can flip our new point
      if(x+1 <= 100) {
        if(M[x+1,y]==1) {flip <- FALSE} #Check right
      }
      if (x >=2 & flip) {
        if(M[x-1,y]==1) {flip <- FALSE} #Check left
      }
      if(y+1<=100 & flip) {
        if(M[x,y+1]==1) {flip<-FALSE} #Check above
      }
      if(y>=2 & flip) {
        if(M[x,y-1] == 1) {flip<-FALSE} #Check below
      }
      if(flip) {M[x,y]<-1}
    } else 
    { M[x,y]<-0 }
    
    #2. now that we have a proposal, we test:
    U <- runif(1)
    
    if(U < (sum(M)^2)/(sum(base)^2)){
     sums <- rbind(sums,(sum(M))^2/10000) 
    } else {
      sums <- rbind(sums,(sum(base))^2/10000)
      M <- base
    }
  }
  return(sums) #this returns a list of values of (f(e))^2
}
```

<p>Now generate a histogram in a similar fashion to part a, generating 100 25,000-step chains:</p>

```{r echo=FALSE}
M4 <- sapply(1:100, function(x) sampler2(25000)) 
qplot(as.numeric(M4), geom="histogram", binwidth=10) + xlab("f(Y)/10000")
```

<p>The new histogram still has the majority of instances trend toward the steady state as the other one did, but since $\frac{\nu(i)}{\nu(j)}$ will actually lead to some moves being rejected, there is a greater concentration of values closer to 0, meaning the number of occupied spaces took longer to increase in the new sampler.</p>

<h5>Problem 3</h5>
</p>(a) Since the order the heads and tails are achieved matters, the space $\Omega$ is found by raising the formula for a permutation of 20 spaces with 11 heads and 9 tails, raised to the 10th power.</p>
$$  \left(\frac{20!}{11!9!}\right)^{10} = 167,960^{10} $$

<p>(b) The formula for $P(\omega | D)$ is the conditional probability of $\omega$ given the results laid out in the problem.</p>
$$ P(\omega | D) = \frac{P(D | \omega)P(\omega)}{P(D)} $$

<p>Logically, since all $\omega$ in $\Omega$ sum to 20 dollars, with each bucket of 20 sequential flips totaling 2 dollars, then $P(D|\omega) = 1$. </p>

<p>The probability of 200 flips resulting in any given $\omega$ is $\frac{1}{2^{200}}$, or one over all possible sequences of 200 flips. </p>

<p>The probability of the given data would be all possible sequences of 200 meeting the conditions, $\Omega$, divided by all possible combinations of 200 flips, or $2^{200}$. </p>

<p>Substituting into the original equation, we get:</p>

$$ P(\omega | D) = \cfrac{1 \times 
                    \cfrac{1}{2^{200}}}{
                    \cfrac{\Omega}{2^{200}}} $$
$$ P(\omega | D) = \frac{1}{\Omega} $$

<p>Thus, the probability of $P(\omega | D) = \frac{1}{\Omega}$, which is uniform over $\Omega$.</p>

<p>(c) The f($\omega$) I decided on to determine the steady state is the amount of money one has after the first 10 flips of a bucket of 20 minus the money collected during the second 10 flips. I chose this function for several reasons:</p>

<ul>
<li>It is never 0, based on the conditions of our problem. If one group of 10 has 5 heads and 5 tails, the other group of 10 cannot have 5 and 5 as well, since there are only 9 tails in each bucket of 20. This is important because the Metropolis-Hastings ratio requires dividing f(j)/f(i).</li>

<li>While the acceptance rate is about 80 percent, it is much lower than other measures I tried, which almost always results in M-H ratios around 1.</li>

<li>It involves all the data in the sequence of 200 in a somewhat meaningful way. Larger test groups, of say 50, would basically be determined by the odd group of 10, because each group of 20 always equals 2 dollars.</li>
</ul>

```{r}
#proposal generating function
flips <- function(chain, num) {
#now we use the current sequence to propose a new sequence by switching up one group of 20 flips
choice <- toString(num)
switch(choice,
       '1'={chain[1:20] <- 1 #knock out all tails
            chain[sample(1:20, 9, replace=F)] <- -1 #place new tails
            },
       '2'={chain[21:40] <- 1
         chain[sample(21:40, 9, replace=F)] <- -1},
       '3'={chain[41:60] <- 1
         chain[sample(41:60, 9, replace=F)] <- -1},
       '4'={chain[61:80] <- 1
         chain[sample(61:80, 9, replace=F)] <- -1},
       '5'={chain[81:100] <- 1
         chain[sample(81:100, 9, replace=F)] <- -1},
       '6'={chain[101:120] <- 1
         chain[sample(101:120, 9, replace=F)] <- -1},
       '7'={chain[121:140] <- 1
         chain[sample(121:140, 9, replace=F)] <- -1},
       '8'={chain[141:160] <- 1
         chain[sample(141:160, 9, replace=F)] <- -1},
       '9'={chain[161:180] <- 1
         chain[sample(161:180, 9, replace=F)] <- -1},
       '10'={chain[181:200] <- 1
         chain[sample(181:200, 9, replace=F)] <- -1},
       stop(chain)
)
return(chain)
}
```

```{r}
#function to sample from all sequences fitting the conditions
omega <- function(N){
results <- data.frame("Step" = 0, "Move" =0, "Diffs" = 0)
chain <- rep(c(rep(1,11), rep(-1,9)),10) #start with a matrix that fits the conditions
num <- trunc(runif(1,1,11)) #now we randomnly pick a bucket of 20 to change
base <- flips(chain, num) #change bucket in base matrix

#moving through the desired number of steps
for ( i in 1:N) {
  num <- trunc(runif(1,1,11)) #choose bucket to change
  proposal <- flips(base, num) #generate a proposal

#f(x) is the difference in the amount of money generated in the first 10 flips of any bucket of 20 minus the amount of money earned in the final 10 flips
proposal.m <- sum(proposal[c(1:10, 21:30, 41:50, 61:70, 81:90, 101:110, 121:130, 141:150, 161:170, 181:190)]) - sum(proposal[c(11:20, 31:40, 51:60, 71:80, 91:100, 111:120, 131:140, 151:160, 171:180, 191:200)])

base.m <- sum(base[c(1:10, 21:30, 41:50, 61:70, 81:90, 101:110, 121:130, 141:150, 161:170, 181:190)]) - sum(base[c(11:20, 31:40, 51:60, 71:80, 91:100, 111:120, 131:140, 151:160, 171:180, 191:200)])

test <- proposal.m/base.m #this results in the proposal being accepted roughly 78% of the time

if(runif(1) < min(1,test, na.rm=TRUE)) {
  base <- proposal
  results <- rbind(results, c(results <- data.frame("Step" = i, "Move" = 1, "Diffs" = proposal.m)))
} else {
  base <- base
  results <- rbind(results, c(results <- data.frame("Step" = i, "Move" = 0, "Diffs" = base.m)))
}
}
return(results)
}
```

<p>Running the sampler 10,000 times shows it reaches a steady state nearly immediately, so I will use 3,000 steps to generate the histogram for part d.</p>

```{r echo=FALSE}
#test runs to show acceptance rate
tests <- omega(10000)
qplot(x = as.numeric(tests$Step), y = as.numeric(tests$Diffs), geom="line") + ylab("Difference in number of tails") + xlab("Step")
```

<p>(d) In order to count the money at each step, we will adjust the output of the omega function:</p>

```{r}
omega2 <- function(N){
  results <- data.frame("Step" = 0, "Round" =0, "Amount" = 0)
  
  #start with a matrix that fits the conditions, but with 1's and 0's instead of 1 and -1 so that any function we do to count tails will always be greater than 0
  chain <- rep(c(rep(1,11), rep(-1,9)),10) 
  
  #now we randomnly pick a bucket of 20 to change
  num <- trunc(runif(1,1,11))
  
  #assign it to the base function
  base <- flips(chain, num)
  
  for ( i in 1:N) {
    
    num <- trunc(runif(1,1,11)) #choose bucket to change
    proposal <- flips(base, num) #generate a proposal
    
    proposal.m <- sum(proposal[c(1:10, 21:30, 41:50, 61:70, 81:90, 101:110, 121:130, 141:150, 161:170, 181:190)]) - sum(proposal[c(11:20, 31:40, 51:60, 71:80, 91:100, 111:120, 131:140, 151:160, 171:180, 191:200)])
    
    base.m <- sum(base[c(1:10, 21:30, 41:50, 61:70, 81:90, 101:110, 121:130, 141:150, 161:170, 181:190)]) - sum(base[c(11:20, 31:40, 51:60, 71:80, 91:100, 111:120, 131:140, 151:160, 171:180, 191:200)])
    
    test <- proposal.m/base.m
    
    #this results in the proposal being accepted roughly 78% of the time
    if(runif(1) < min(1,test, na.rm=TRUE)) {
      base <- proposal
      total <- cumsum(as.numeric(base))
      top.dollar <- which.max(total)
      results <- rbind(results, c(results <- data.frame("Step" = i, "Round" = top.dollar, "Amount" = total[top.dollar])))
    } else {
      base <- base
      total <- cumsum(as.numeric(base))
      top.dollar <- which.max(total)
      results <- rbind(results, c(results <- data.frame("Step" = i, "Round" = top.dollar, "Amount" = total[top.dollar])))
    }
  }
  return(results)
}
```

<p>And the resulting histograms for 3,000 steps:</p>

```{r echo =FALSE}
tests2 <- omega2(3000)
qplot(as.numeric(tests2$Round[2:3000]), geom="histogram", binwidth = 1, fill=..count.., main = "Histogram for T") + xlab("Round Max Achieved")
qplot(as.numeric(tests2$Amount[2:3000]), geom="histogram", binwidth = 1, fill = ..count.., main = "Histogram for Z") + xlab("Highest Amount")
```
